---
title             : "A Time-varying Computational Model of Learning to Read Printed Words Aloud"
bibliography      : ["references.bib"]
documentclass     : "apa6"
figsintext        : yes
figurelist        : no
tablelist         : yes
footnotelist      : no
lineno            : no
lang              : "en-US"
output            :
  papaja::apa6_pdf :
    toc               : false
    toc_depth         : 3
    indent            : true

header-includes:
  - \geometry{left=3cm, right=3cm, top=3cm, bottom=3cm}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \PassOptionsToPackage{x11names}{xcolor}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage[normalem]{ulem}
  - \usepackage{pdfpages}
  - \interfootnotelinepenalty=10000
  - \setcounter{page}{-2}
  - \AtBeginDocument{\let\maketitle\relax}
  - \pagestyle{plain}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[R]{\thepage}
  # for title page:
  - \usepackage{setspace}
  - \usepackage{afterpage}
  - \usepackage{everypage}

---

\thispagestyle{empty}
\newpage

\begin{center}
\thispagestyle{empty}
\vspace{5cm}
A Time-varying Computational Model of Learning to Read Printed Words Aloud \\
\vfil
by \\  
Matthew James Cooper Borkenhagen \\
\vfil
A dissertation submitted in partial fulfillment of \\
the requirements for the degree of \\
\vfil
Doctor of Philosophy \\
(Psychology) \\
\vfil
at the \\
\textsc{University of Wisconsin - Madison} \\
2023 \\
\vfil

\vfil
\end{center}
\begin{singlespace}
Date of final oral examination: January 21st, 2022 \newline
The dissertation is approved by the following members of the Final Oral Committee: \newline
\hspace*{2cm} Mark S. Seidenberg, Professor, Psychology \newline
\hspace*{2cm} Martha Alibali, Professor, Psychology \newline
\hspace*{2cm} Maryellen MacDonald, Professor, Psychology \newline
\hspace*{2cm} Robert Nowak, Professor, Electrical and Computer Engineering \newline
\hspace*{2cm} Timothy T. Rogers, Professor, Psychology
\end{singlespace} 
\clearpage

\newpage
\thispagestyle{empty}

\vspace*{\fill}

\centerline{\copyright \; Copyright by Matthew J. Cooper Borkenhagen - 2023}
\centerline{All Rights Reserved}

\newpage


\pagenumbering{roman}

\newpage

\centering
\raggedright
\newpage
\tableofcontents


```{r global_options, include = FALSE, eval = F, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      cache = FALSE,
                      echo = FALSE,
                      out.extra = "",
                      fig.pos = "t!",  sanitize = TRUE)
options(knitr.table.format = "latex")
options(knitr.kable.NA = '')
options(scipen = 1000)
#knitr::knit_hooks$set(inline = function(x) {if(!is.numeric(x)){x}else{prettyNum(round(x,2), big.mark=",")}})
#knit_hooks$set(inline = function(x) {prettyNum(x, big.mark=",")})
```


```{r loads, echo=FALSE, warning=FALSE, include=FALSE}

# requirements
packages = read.csv('requirements.txt', stringsAsFactors = F)[[1]]
lapply(packages, require, ch = T)

source('scripts/utilities.R')
# we reference some ch3 data here in chapter 2 so we will compile all data here
# note that source scripts are in ./data/
source('scripts/data.R')
```


\newpage

```{r acknowledgements, child='acknowledgements.Rmd'}
```

\newpage
\pagenumbering{arabic}
\setlength\parindent{25pt}
```{r child='1_chapter.Rmd'}
```

```{r child='2_chapter.Rmd'}
```

```{r child='3_chapter.Rmd'}
```

```{r child='4_chapter.Rmd'}
```

\newpage

```{r child='5_appendixA.Rmd'}
```


\newpage

```{r child='6_appendixB.Rmd'}
```


\newpage

```{r child='7_appendixC.Rmd'}
```

\pagebreak


# References

