---
title             : 'Appendix 1'
shorttitle        : 'Appendix 1'
output            : papaja::apa6_pdf
#classoption       : "man, donotrepeattitle"

header-includes:
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\singlespacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1.5}}
  - \pagenumbering{gobble}
#  - \AtBeginDocument{\let\maketitle\relax}

---


```{r echo=FALSE}
require(tidyverse)
require(papaja)
```

## Phonological representations
Appendix 1 contains information about the phonological representations used for training the computational models described in this work. The features used were primarily taken from those in Harm & Seidenberg (1999) and Harm & Seidenberg (2004). Nodes were added for stress, where in the tables the columns "1ary" refers to primary stress and "2ary" refers to secondary stress. A phoneme that has both of those units set to zero would be unstressed, as is the case with unstressed vowels and consonants. There are also units for "start" and "end", which are only used by the start and end of (phonological) word segments. For those segments all units are zero and either of these two units is set to on. As is noted at the bottom of each table, consonants are also represented with all the vowel features and vowel features are represented with all consonant features. In any given table only the relevant features are shown in order to allow the tables to fit on the page. Any feature not shown for a particular phoneme should be assumed to be set to zero. Table 6 shows the special representations for the null segment ("_"), start of word segment ("#"), and the end of word segment ("%"), all of which have special purposes in training and testing. Note, however, that the null segment is never used in training, though it is sometimes produced in testing and is referred to as an "indeterminate segment" when described in Chapter 3.

\newpage

```{r CMUPhonKeyVowels, echo=FALSE}

read_csv('data/cmu_phoneme_key.csv') %>% 
  filter(Category == 'V') %>% 
  select(Phon, Example) %>% 
  apa_table(caption = 'Symbolic Representations and Examples of Vowel Phonemes Used', note = 'Phonemes are referred to with the abbreviation "phon" in this appendix', small = TRUE, longtable = TRUE)


```

\newpage

```{r CMUPhonKeyConsonantss, echo=FALSE}

read_csv('data/cmu_phoneme_key.csv') %>% 
  filter(Category == 'C') %>% 
  select(Phon, Example) %>% 
  apa_table(caption = 'Symbolic Representations and Examples of Vowel Phonemes Used', small = TRUE, longtable = TRUE)


```


```{r consonants, echo=FALSE}
read_csv('data/phonreps_consonants.csv') %>% 
  mutate_if(is.numeric, as.character) %>% 
  filter(Phon %nin% c('DX', 'NX', 'ZH', '_', '#', '%')) %>% 
  apa_table(caption = 'Phonological Representations for Consonants', note = '"Alveo." = "alveolar", "Fric." = "fricative", "Affric." = "affricate", and "Rhot." = "rhotic". In addition to the features (columns) shown, these phonemes are also represented with the features depicted for the vowels shown in other tables in this appendix, but with those features set to 0.', landscape = TRUE, small = TRUE, longtable = TRUE)
```





```{r UnstressedVowels, echo=FALSE}
read_csv('data/phonreps_vowels1.csv') %>% 
  mutate_if(is.numeric, as.character) %>% 
  filter(Phon %nin% c('_', '#', '%')) %>% 
  apa_table(caption = 'Phonological Representations for Vowels Recieving No Stress', note = '"Phon" = "Phoneme", "Cent." = "Central", "C-mid" = "Close-mid", "O-mid" = "Open-mid", "N-open" = "Near-open", and "Rnd." = "Round". "y_" is equivalent to "Post-y", and ""w_" is equivalent to "Post-w". In addition to the features (columns) shown, these phonemes are also represented with the features depicted for the consonants provided Table 2 in this appendix, but with those features set to 0. They are not depicted here for simplicity. The features "start" and "end" are used only for identifying the special segments "#" (start of word) and "%" (end of word) respectively, which can be found in the table on special segments in this appendix.', landscape = TRUE, small = TRUE, longtable = TRUE)
```


```{r PrimaryStressedVowels, echo=FALSE}
read_csv('data/phonreps_vowels2.csv') %>% 
  mutate_if(is.numeric, as.character) %>% 
  apa_table(caption = 'Phonological Representations for Vowels Recieving Primary Stress', note = '"Phon" = "Phoneme", "Cent." = "Central", "C-mid" = "Close-mid", "O-mid" = "Open-mid", "N-open" = "Near-open", and "Rnd." = "Round". "y_" is equivalent to "Post-y", and ""w_" is equivalent to "Post-w". In addition to the features (columns) shown, these phonemes are also represented with the features depicted for the consonants provided Table 2 in this appendix, but with those features set to 0. They are not depicted here for simplicity. The features "start" and "end" are used only for identifying the special segments "#" (start of word) and "%" (end of word) respectively, which can be found in the table on special segments in this appendix.', landscape = TRUE, small = TRUE, longtable = TRUE)
```

```{r SecondaryStressedVowels, echo=FALSE}
read_csv('data/phonreps_vowels2.csv') %>% 
  mutate_if(is.numeric, as.character) %>% 
  apa_table(caption = 'Phonological Representations for Vowels Recieving Secondary Stress', note = '"Phon" = "Phoneme", "Cent." = "Central", "C-mid" = "Close-mid", "O-mid" = "Open-mid", "N-open" = "Near-open", and "Rnd." = "Round". "y_" is equivalent to "Post-y", and ""w_" is equivalent to "Post-w". In addition to the features (columns) shown, these phonemes are also represented with the features depicted for the consonants provided Table 2 in this appendix, but with those features set to 0. They are not depicted here for simplicity. The features "start" and "end" are used only for identifying the special segments "#" (start of word) and "%" (end of word) respectively, which can be found in the table on special segments in this appendix.', landscape = TRUE, small = TRUE, longtable = TRUE)
```


```{r SpecialSegments, echo=FALSE}
read_csv('data/phonreps_vowels1.csv') %>% 
  mutate_if(is.numeric, as.character) %>% 
  filter(Phon %in% c('_', '#', '%')) %>% 
  apa_table(caption = 'Phonological Representations for Special Segments', note = 'Representations were adapted from Harm & Seidenberg (1999). In addition to the features (columns) shown, these phonemes are also represented with the features depicted for the consonants provided Table 2 in this appendix, but with those features set to 0. They are not depicted here for simplicity. Note that the null "_" segment has all units set to off, but is never used in training except for the feedforward models reported.', landscape = TRUE, small = TRUE, longtable = TRUE)
```
